Detection and Classification of Supernova Gravitational Waves Signals: A Deep Learning Approach

### 2019-12-09 main.tex
change the text in the test waveform section
Made a minor change to see if I can commit
Abouthalf wayt through the draft with my comments

% avoids incorrect hyphenation, added Nov/08 by SSR
%\hyphenation{ALPGEN}
%\hyphenation{EVTGEN}
%\hyphenation{PYTHIA}
%\renewcommand{\bibfont}{\normalfont\small}

@@

\author{Authors$^{12}$}
\affiliation{
$^1$Department of Applied Physics, Fukuoka University, Nanakuma 8-19-1, Fukuoka 814-0180, Japan\\
$^2$SUPA, School of Physics and Astronomy, University of Glasgow, Glasgow G12 8QQ, UK}
%$^2$Caltech CaRT, Pasadena, California 91125, USA}
%\date{\today}
%\date{\commitDATE\\\mbox{\small \commitID}\\\mbox{\dcc}}

\begin{abstract}
We demonstrate the application of a convolutional neural network 
to the gravitational wave signals from core collapse supernovae.
Using simulated time series of gravitational wave detectors, 
we show that a convolutional neural network can be used to detect and classify, based on the explosion mechanisms,  
the gravitational wave signals buried in noise.
For the waveforms used in the training of the convolutional neural network, 
our results suggest that a network of advanced LIGO, advanced VIRGO and KAGRA, or a network of LIGO A+, advanced VIRGO and KAGRA
is likely to detect a magnetorotational core collapse supernovae within the Large and Small Magellanic Clouds, or 
a Galactic event if the explosion mechanism is the neutrino-driven mechanism.
By testing the convolutional neural network with waveforms not used for training, 
we show that the true alarm probabilities are $52\%$ and $83\%$ at $60$kpc for waveforms $\text{R3E1AC}$ and $\text{R4E1FC\_L}$.
For waveforms $\text{s}20$ and $\text{SFHx}$ at $10$ kpc, the true alarm probabilities are $70\%$ and $93\%$ respectively. All at false alarm probability equal to $0.1$.
\end{abstract}
\pacs{}
\maketitle
\acrodef{GW}[GW]{gravitational wave}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{NSBH}[NSBH]{neutron star black hole}
\acrodef{EM}[EM]{electromagnetic}
\acrodef{CBC}[CBC]{compact binary coalescence}
\acrodef{CNN}[CNN]{Convolutional neural network}
\acrodef{CCSN}[CCSN]{core collapse supernova}
\acrodefplural{CCSN}[CCSNe]{core collapse supernovae}
\acrodef{ROC}[ROC]{receiver operator characteristic}
\acrodef{TAP}[TAP]{true alarm probability}
\acrodef{FAP}[FAP]{false alarm probability}
\acrodef{aLIGO}[aLIGO]{advanced LIGO}
\acrodef{AdVirgo}[AdVirgo]{advanced VIRGO}

\section{Introduction}
Since 2015 when LIGO has made the first direct observation of \acp{GW} from the merger of a binary black hole\cite{abbott2016observation},
there have been numerous observations of \acp{GW} from similar systems in the observation runs of LIGO 
and VIRGO \cite{abbott2016gw151226, abbott2017gw170608, abbott2017gw170814}.
These discoveries have been a crucial milestone in \ac{GW} astronomy and have opened up a new window on the sky.
More recently, LIGO and VIRGO have observed the \acp{GW} from a binary neutron star merger\cite{abbott2017gw170817, abbott2017gravitational, abbott2017multi}.
For this event, the \acp{GW} and the associated gamma-ray burst were observed simultaneously. 
Other counterparts across the electromagnetic spectrum were also observed by later follow-up observations \cite{abbott2017multi}.
In the near future, many more observations of \acp{GW} from similar compact binary coalescence systems can be expected
as KAGRA starts joint observations with LIGO and VIRGO \cite{aso2013interferometer, somiya2012detector, abbott2018prospects}.

In addition to compact binary coalescences, massive stars with $10-100 \text{M}_\odot $ at zeros-age main sequence ending their lives by becoming 
%with core of $10-100 \text{M}_\odot $ prior to collapse 
\acp{CCSN} are also considered to be potential sources to the second generation detectors such as the \ac{aLIGO} \cite{aasi2015advanced}, 
\ac{AdVirgo} \cite{acernese2014advanced} and KAGRA interferometers \cite{aso2013interferometer, gossan2016observing, abbott2016first}.
It is currently not entirely clear to astronomers how such a massive star becomes supernova. 
The basic theory of the explosion, confirmed by the neutrino events observed from SN1987A\cite{sato1987analysis}, goes as the following.
A massive star at the final stage of its life will form core that is composed of iron nuclei after it has burned all its stellar fuel via nuclear reaction.
The iron core is supported by the pressure of relativistic degenerate electrons.
If the mass of the core exceeds the effective Chandrasekhar mass\cite{baron1990effect, bethe1990supernova}, core collapse will ensue and continue until it reaches nuclear density.
The nuclear equation of state will then stiffens by the strong nuclear force above the nuclear density and stops 
the core collapse.
The inner core will bounce back and a shock wave will be sent through the infalling matter.
By losing energy to the dissociation of the iron nuclei and to neutrino cooling, the shock wave will stall.
For the star to become a supernova, the shock wave will need to be revived\cite{o2011black}.  
The mechanism via which the shock wave is revived and causes the explosion has been the subject of intense study.
However, it still remains an unsolved problem.

There exist two most popular theories, the neutrino-driven mechanism\cite{bethe1985revival, bethe1990supernova} and the magnetorotational mechanism\cite{janka2012explosion, kotake2012core, mezzacappa2014two}.
For supernova progenitors with core rotation too slow to affect the dynamics\cite{takiwaki2016three, summa2018rotation}, the neutrino-driven mechanism is believed to be the active mechanism.
The majority of the observed \acp{CCSN} can be explained by the neutrino mechanism\cite{bruenn2016development}.
The neutrino mechanism\cite{bethe1985revival, janka2007theory} suggests that about $5-10\%$ of the outgoing neutrino luminosity is stored below the shock, 
which causes turbulence to occur and thermal pressure to increase.
The stalled shock can be revived by their combined effects\cite{couch2015role}. 
Producing a \ac{CCSN} via the neutrino mechanism may also require convection and the standing accretion shock instability\cite{blondin2003stability}.
On the other hand, the magnetorotational mechanism requires rapid core spin and strong magnetic field\cite{leblanc1970numerical, burrows2007simulations, takiwaki2009special, moiseenko2006magnetorotational,mosta2014magnetorotational}. 
Together, they may produce an outflow that may cause the most energetic \acp{CCSN} observed.
The magnetorotational mechanism may be able to explain the extreme hypernovae and the observed long gamma-ray bursts\cite{woosley2006progenitor, yoon2005evolution, de2013rotation}.

Correctly classifying the \ac{GW} from a \ac{CCSN} is important in understanding the explosion mechanism.
%Electromagnetic observations provide only indirect informations about the mechanism of a \ac{CCSN} 
%because photons interact strongly with matter and only information from the optically thin regions far from the central core could reach us. 
%However, 
As \acp{GW} are emitted in the central core of a \ac{CCSN}, 
they are likely to carry direct information of the \ac{CCSN} and therefore provide a probe of the explosion mechanism that produces them.
In \ac{GW} astronomy, when it comes to the search for signals from compact binary coalescences, the established routine is matched filtering.
However, as the emission process of the \acp{GW} from \acp{CCSN} is affected by turbulence in the post-bounce and is stochastic in nature,  
the signal evolution cannot be predicted robustly\cite{ott2009gravitational, kotake2013multiple, kotake2009stochastic}.
This in turn prevents the match filtering routine to be applied to \acp{CCSN}.
%Although, Neutrino radiation-hydrodynamics simulations have been an ongoing effort to decode and understand explosion mechanism of \acp{CCSN} and the \acp{GW} they produce\cite{ott2009gravitational}.
Methods and algorithms have been developed for the detection and classification of signals from \acp{CCSN}. 
For example, a method known as principle component analysis has been developed\cite{heng2009rotating, rover2009bayesian, powell2015classification, powell2017classification, suvorova2019reconstructing}.
This method creates a set of component basis vectors from a set of \ac{CCSN} waveforms of a particular mechanism
that represent the common features of the waveforms of that mechanism.
There have been other approaches developed in the literature such as Bayesian inference\cite{rover2009bayesian}, Bayesian model selection\cite{logue2012inferring}
, multivariate regression model\cite{engels2014multivariate}, maximum entropy\cite{summerscales2008maximum}, 
maximum likelihood\cite{gursel1989near} and Tikhonov regularization scheme\cite{rakhmanov2006rank, hayama2007coherent}.

In recent years, the field of machine learning and its sub-field, deep learning, have been rapidly developing because of 
its potentials in many fields \cite{krizhevsky2012imagenet, NIPS2014_5423, simonyan2014very, chen2014semantic, zeiler2014visualizing, szegedy2015going}.
For example, deep learning have been successfully applied to fields including medical diagnosis\cite{kononenko2001machine}, object detection\cite{redmon2016you}, image recognition/processing/generation\cite{he2016deep, krizhevsky2012imagenet, zhang2016colorful, karpathy2015deep},    and  language processing\cite{lample2016neural}.
In \ac{GW} astronomy, deep learning has mostly been applied to both the identifications of glitch\cite{mukund2017transient, zevin2017gravity, george2017deep, gabbard2018matching} and signals\cite{george2018deep, astone2018new}. 
\ac{CNN} is a deep learning algorithm that has the advantage of capturing spatial and temporal features of the input data.
Another advantage of the use of a \ac{CNN} in the detection of a signal is that a \ac{CNN} is relatively computationally cheap compared to other more traditional methods.
This is because the heavy computational work is usually done during the training stage of a \ac{CNN} prior to its actual application\cite{goodfellow2016deep}.

In this work, we demonstrate a \ac{CNN} can be applied to the detection of the \acp{GW} from \acp{CCSN}
and the classification of their explosion mechanisms.
To train our \ac{CNN}, we use simulated \ac{CCSN} waveforms from a number of studies and simlate sources 
at a range of distances from $10$ to $200$ kpc for two networks of four detectors.
The first network consists of \ac{aLIGO}, \ac{AdVirgo} and KAGRA.
For the second network, we still include the detectors of \ac{AdVirgo} and KAGRA,  
but replace the two detectors of \ac{aLIGO} with 
a modest set of planned upgrade version of them - LIGO A+ in Hanford and Livingston\cite{miller2015prospects, LIGOW}.
We will then apply the \ac{CNN} to waveforms excluded from the training session to show that the \ac{CNN} 
is not only able to identify the waveforms it familiarised itself with during training, but also new waveforms.

The remaining of this paper is constructed as follows. 
In section \ref{sec:CNN}, we will present a brief explanation of the concept of \ac{CNN} as well as the \ac{CNN} we used for this work.
In section \ref{sec:spwf}, we introduce the waveforms we use for the training of the \ac{CNN} in this work. We will also discuss the procedure with which we generated the data for the \ac{CNN}.
The results will be shown in sections \ref{sec:result} and \ref{sec:unseen}, followed by a conclusion in section \ref{sec:conclusion}.
%\section{Supernova waveform}\label{sec:spwf}

\section{Convolutional Neural network}\label{sec:CNN}
A \ac{CNN} is a computational processing system that is composed of interconnected layers of computational nodes\cite{o2015introduction}.
The nodes are known as neurons and are associated with an activation function.
The activation function can perform an elementwise nonlinear operation to the input of the layer. 
In a \ac{CNN}, there are three types of layers: convolutional layer, max-pooling layer, and fully connected layer\cite{o2015introduction}.
Convolutional layer performs the mathematical operation of convolution between the weights of the layer's neurons and the input to that layer.
Max-pooling layer is just a down-sampling layer that down samples the input along its dimensionality.
It can reduce the computational cost by decreasing the number of parameters of the \ac{CNN}.
Fully connected layer is a layer that connects every neuron in its layer to every neuron of its immediate previous and next layer.
In the case of detection and classification, fully connected layers are used to compute class scores.

When multiple layers of these three types are stacked and connected one after the other, 
a \ac{CNN} has been formed, where the output of each layer is the input of the next layer.
How these layers are connected in a \ac{CNN} is known as the architecture of the \ac{CNN}.
It describes the structure of a \ac{CNN} and the number of neurons in each layer. % (width and depth) 
In general, the first layer in a \ac{CNN}, also known as the input layer, is often a convolutional layer, 
while the last layer or the output layer is often a fully connected layer with an associated loss function. 
The architecture of the rest of the \ac{CNN} should depend on the specific task that the \ac{CNN} is trained to solve. 
An over-complicated model with too many trainable parameters 
is easier to result in overfit and harder to train, 
while a too-simple \ac{CNN} will have a hard time capturing the feature inherent to the input.
In addition, the number of layers and neurons of each layer in a \ac{CNN} are known as hyperparameters.
Other hyperparameters include the parameters of max-pooling layers, type of activation functions, 
learning rate of the \ac{CNN}, and the application of specific deep learning techniques if there is any.
The optimal combination of hyperparameters and the architecture is sought by trail and error, and fine tunning.

During the training stage, the weights of the neurons in a \ac{CNN} are updated using an algorithm called back propagation\cite{lecun1988theoretical}.
The output of the \ac{CNN} is used as an input to the loss function associated with the output layer. 
The back propagation algorithm will then compute the gradient of the loss function 
which is used to adjust the values of the weights of the neurons in each layer and minimise the loss function.
When the loss function is minimised, the \ac{CNN} will classify its input into the correct category with the highest confidence. 
The process of achieving the minimisation of the loss function during the training stage is referred to as learning.

In this work, we employ a \ac{CNN} of $8$ convolutional layers, $3$ max-pooling layers, and $3$ fully connected layers.
The exact architecture of the \ac{CNN} is shown in Table \ref{table:architecture} and illustrated in Figure \ref{fig:CNN}.
\begin{table}[]
\centering
\begin{threeparttable}
\caption{The architecture of the \ac{CNN}}
\label{table:architecture}
\begin{tabular}{ccccc}
\\
\toprule
Layer & Type      & Neurons  & Filter size & Act. Fun \\ \hline
1     & Conv      & 11        & 32          & Elu      \\ 
2     & Max-pool  &          & 8          &          \\
3     & Conv      & 11        & 8          & Elu      \\
4     & Max-pool  &          & 6           &          \\
5     & Conv      & 11       & 6           & Elu      \\
6     & Conv      & 11       & 4           & Elu      \\
7     & Conv      & 13       & 4           & Elu      \\
8     & Conv      & 13       & 4           & Elu      \\
9     & Conv      & 13       & 4           & Elu      \\
10    & Conv      & 13       & 4           & Elu      \\
11    & Max-pool  &          & 2           &          \\
12    & Fully-con & 64(50\%) &             & Elu      \\
13    & Fully-con & 32(50\%) &             & Elu      \\
14    & Fully-con & 3        &             & Softmax      \\ 
\hline
\hline
\end{tabular}
\begin{tablenotes}
\setlength\labelsep{0pt}
\normalfont{
\item The architecture of the \ac{CNN} used in this work for the purpose of distinguishing supernova signal mechanisms and background noise.
In the table, Conv means convolutional neural layer, Max-pool max-pooling layer, and Fully-con fully-connected layers.
Neuron and Act. Fun mean the number of neurons and the activation function for the layer respectively. 
The numbers in the bracket for the fully-connected layers are the number used for drop-out.}
\end{tablenotes}
\end{threeparttable}
\end{table}
\begin{figure}
\includegraphics[width=0.48\textwidth]{CNN2.png}
\caption{An illustration of the architecture of the \ac{CNN} used in this paper for the detection and classification of \ac{CCSN} signals in noisy data.
The \ac{CNN} consists of $8$ convolutional layers, $3$ max-pooling layers and $3$ fully connected layer including the output layer.
The input layer takes the simulated time series of the detectors as input, feeding through the \ac{CNN}. The \ac{CNN} will output three probabilities at the last layer.
The numbers above or below each layer indicate the kernal size of the layer. For example, the first convolutional layer has $11$ filters, 
each of which is $1$ by $32$ in size. 
The elements in the figure are not to scale.
\label{fig:CNN}}
\end{figure}
Since the problem we are trying to solve is a problem of multi-class classification, 
the loss function employed for this work is categorical cross entropy\cite{abadi2016tensorflow}, given by
\begin{equation}\label{eq:cce}
 L(y,\hat{y}) = -\sum^M_{j=1}\sum^C_{i=1}(y_{ij}log(\hat{y}_{ij})),
\end{equation}
where $C$ is the number of the classes and $M$ is the number of the training samples.
For the $j^{\text{th}}$ sample and the $i^{th}$ class, $y_{ij}$ is the correponding class value. It is equal to $1$ for the true class and $0$ otherwise.
Similarly, $\hat{y}_{ij}$ is the predicted probability from the \ac{CNN} for the $i^{th}$ class and the $j^{\text{th}}$ sample.

\section{Data}\label{sec:spwf}
We establish a \ac{CNN} for the purpose of distinguishing detector time series among three classes, 
i.e., magnetorotational signals + background noise, neutrino-driven signals + background noise, and pure background noise. 
For this purpose, it is necessary to prepare training, validation and testing data of these three classes.
The training data is used for tuning the weights of the neurons in the layers in the \ac{CNN}, 
while validating data is to verify that the \ac{CNN} is learning the features inherent to the data and testing data is to test the performance of the trained \ac{CNN}.

In general, the input of a \ac{CNN} is numerical data.
In our case, a data sample is a set of simulated time series stacked together as a $k \times p$ matrix 
where $k$ is the number of detectors and $p$ the number of elements in the time series.
To this end, we use simulated waveforms in the literature.
The magnetorotational \ac{CCSN} signals are taken from\cite{abdikamalov2014measuring, dimmelmeier2008gravitational, richers2017equation}. 
The simulations in \cite{abdikamalov2014measuring} were focused on the dependence of the waveforms on the angular momentum distrubion of the progenitors and generated 92 waveforms.
In \cite{dimmelmeier2008gravitational}, 136 waveforms were generated to investigate a variety of rotation rates and masses of the progenitors.
The simulations in \cite{richers2017equation} covered a parameter space of $18$ different equations of state and $98$ rotation profiles for a progenitor of $12M_{\odot}$ generating in total $1824$ waveforms.
All the simulations from the studies were 2D.

>>>>

% avoids incorrect hyphenation, added Nov/08 by SSR
%\hyphenation{ALPGEN}
%\hyphenation{EVTGEN}
%\hyphenation{PYTHIA}
%\renewcommand{\bibfont}{\normalfont\small}
% test comment

@@

\author{Authors$^{12}$\cm{fix}}
\affiliation{
$^1$Department of Applied Physics, Fukuoka University, Nanakuma 8-19-1, Fukuoka 814-0180, Japan\\
$^2$SUPA, School of Physics and Astronomy, University of Glasgow, Glasgow G12 8QQ, UK}
%$^2$Caltech CaRT, Pasadena, California 91125, USA}
%\date{\today}
%\date{\commitDATE\\\mbox{\small \commitID}\\\mbox{\dcc}}

\begin{abstract}
We demonstrate the application of a convolutional neural network to the
gravitational wave signals from core collapse supernovae. Using simulated time
series of gravitational wave detectors, we show that based on the explosion
mechanisms, a convolutional neural network can be used to detect and classify
the gravitational wave signals buried in noise. For the waveforms used in the
training of the convolutional neural network, our results suggest that a
network of advanced LIGO, advanced VIRGO and KAGRA, or a network of LIGO A+,
advanced VIRGO and KAGRA is likely to detect a magnetorotational core collapse
supernovae within the Large and Small Magellanic Clouds, or a Galactic event if
the explosion mechanism is the neutrino-driven mechanism.  By testing the
convolutional neural network with waveforms not used for training, we show that
the true alarm probabilities are $52\%$ and $83\%$ at $60$kpc for waveforms
$\text{R3E1AC}$ and $\text{R4E1FC\_L}$.  For waveforms $\text{s}20$ and
$\text{SFHx}$ at $10$ kpc, the true alarm probabilities are $70\%$ and $93\%$
respectively. All at false alarm probability equal to $0.1$\cm{why is this nopt
also written as a percentage? Should it be 10\%?}.  \end{abstract}

\pacs{}
\maketitle
\acrodef{GW}[GW]{gravitational wave}
\acrodef{BNS}[BNS]{binary neutron star}
\acrodef{BBH}[BBH]{binary black hole}
\acrodef{NSBH}[NSBH]{neutron star black hole}
\acrodef{EM}[EM]{electromagnetic}
\acrodef{CBC}[CBC]{compact binary coalescence}
\acrodef{CNN}[CNN]{Convolutional neural network}
\acrodef{CCSN}[CCSN]{core collapse supernova}
\acrodefplural{CCSN}[CCSNe]{core collapse supernovae}
\acrodef{ROC}[ROC]{receiver operator characteristic}
\acrodef{TAP}[TAP]{true alarm probability}
\acrodef{FAP}[FAP]{false alarm probability}
\acrodef{aLIGO}[aLIGO]{advanced LIGO}
\acrodef{AdVirgo}[AdVirgo]{advanced VIRGO}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%
% general background on GWs
%
Since 2015 when LIGO made the first direct observation of \acp{GW} from the
merger of a binary black hole~\cite{abbott2016observation}, there have been
numerous observations of \acp{GW} from similar systems in the subsequent
observation runs of LIGO and VIRGO~\cite{abbott2016gw151226,
abbott2017gw170608, abbott2017gw170814}. These discoveries represent a crucial
milestone in \ac{GW} astronomy and have opened up a new window on the universe.
More recently, LIGO and VIRGO have observed the \acp{GW} from a binary neutron
star merger~\cite{abbott2017gw170817, abbott2017gravitational, abbott2017multi}
where the \acp{GW} and an associated gamma-ray burst were observed
simultaneously. Other counterparts across the electromagnetic spectrum were
also observed by later follow-up observations~\cite{abbott2017multi}. In the
near future, many more observations of \acp{GW} from similar \ac{CBC} systems
can be expected as KAGRA starts joint observations with LIGO and
VIRGO~\cite{aso2013interferometer, somiya2012detector, abbott2018prospects}.

%
% intro to core-collapse supernovae
%
In addition to \acp{CBC}, massive stars with $10-100 \text{M}_\odot $ at
zeros-age main sequence ending their lives by becoming 
%with core of $10-100 \text{M}_\odot $ prior to collapse 
\acp{CCSN} are also considered to be potential sources to the second generation
detectors such as \ac{aLIGO}~\cite{aasi2015advanced},
\ac{AdVirgo}~\cite{acernese2014advanced} and KAGRA
interferometers~\cite{aso2013interferometer, gossan2016observing,
abbott2016first}. It is currently not entirely clear to astronomers how such
massive stars become supernovae. The basic theory of the explosion, confirmed
by the neutrino events observed from SN1987A~\cite{sato1987analysis}, begins
with a massive star at the final stage of its life forming a core that is
composed of iron nuclei after it has burned all its stellar fuel via fusion
reactions. The iron core is supported by the pressure of relativistic
degenerate electrons and if the mass of the core exceeds the effective
Chandrasekhar mass~\cite{baron1990effect, bethe1990supernova}, core collapse
will ensue and continue until the core reaches nuclear density. The nuclear
equation of state will then stiffens by the strong nuclear force above the
nuclear density and stops the core collapse. The inner core will bounce back
and a shock wave will be sent through the infalling matter.  By losing energy
to the dissociation of the iron nuclei and to neutrino cooling, the shock wave
will stall. For the star to become a supernova, the shock wave will need to be
revived~\cite{o2011black}. However the mechanism via which this occurs and the
supernova explosion is caused has been the subject of intense study and remains
an unsolved problem.

%
% discussion of the CCSN mechanisms
%
There exist two most popular theories~\cm{theories for what?}, the
neutrino-driven mechanism~\cite{bethe1985revival, bethe1990supernova} and the
magnetorotational mechanism~\cite{janka2012explosion, kotake2012core,
mezzacappa2014two}. For supernova progenitors with core rotation too slow to
affect the dynamics~\cite{takiwaki2016three, summa2018rotation}, the
neutrino-driven mechanism is believed to be the active mechanism. The majority
of the observed \acp{CCSN} can be explained by the neutrino
mechanism~\cite{bruenn2016development}~\cm{why?}. The neutrino
mechanism~\cite{bethe1985revival, janka2007theory} suggests that approximately
$5-10\%$ of the outgoing neutrino luminosity is stored below the shock, which
causes turbulence to occur and thermal pressure to increase. The stalled shock
can be revived by their combined effects~\cite{couch2015role}. Producing a
\ac{CCSN} via the neutrino mechanism may also require convection and the
standing accretion shock instability~\cite{blondin2003stability}. On the other
hand, the magnetorotational mechanism requires rapid core spin and a strong
magnetic field~\cite{leblanc1970numerical, burrows2007simulations,
takiwaki2009special,
moiseenko2006magnetorotational,mosta2014magnetorotational}. Together, they may
produce an outflow that may cause the most energetic \acp{CCSN}
observed~\cm{are you referring to a real observed CCSN, if so you should
provide a reference.}. The magnetorotational mechanism may \cm{also?} be able
to explain the extreme hypernovae and the observed long gamma-ray
bursts\cite{woosley2006progenitor, yoon2005evolution, de2013rotation}.

%
% GWS from CCSN and the existing literature
%
Correctly classifying the \ac{GW} signals from a \ac{CCSN} is important in
understanding the explosion mechanism.
%Electromagnetic observations provide only indirect informations about the
%mechanism of a \ac{CCSN} because photons interact strongly with matter and
%only information from the optically thin regions far from the central core
%could reach us.  However, 
As \acp{GW} are emitted\cm{generated?} in the central core of a \ac{CCSN}, they
are likely to carry direct information of the \ac{CCSN} and therefore provide a
probe of the explosion mechanism that produces them. In \ac{GW} astronomy, for
\ac{CBC} events the established method for searching for signals is
matched-filtering~\cm{give ref for pycbc and gstlal}. However, since the
emission process of the \acp{GW} from \acp{CCSN} is affected by turbulence in
the post-bounce and is \cm{expected to be?} stochastic in nature,  the signal
evolution cannot be predicted robustly~\cite{ott2009gravitational,
kotake2013multiple, kotake2009stochastic}.  This in turn prevents
matched-filtering from being applied to \acp{CCSN}.
%Although, Neutrino radiation-hydrodynamics simulations have been an ongoing
%effort to decode and understand explosion mechanism of \acp{CCSN} and the
%\acp{GW} they produce\cite{ott2009gravitational}.
Methods and algorithms have been developed for the detection and classification
of signals from \acp{CCSN}.  For example, a method known as principle component
analysis has been developed~\cite{heng2009rotating, rover2009bayesian,
powell2015classification, powell2017classification,
suvorova2019reconstructing}. This method creates a set of component basis
vectors from a \cm{larger?} set of \ac{CCSN} waveforms belonging to a
particular mechanism where the basis vectors represent the common features of
those waveforms.  There have been other approaches developed in the literature
such as Bayesian inference~\cite{rover2009bayesian}, Bayesian model
selection~\cite{logue2012inferring} , multivariate regression
modelling~\cite{engels2014multivariate}, maximum
entropy~\cite{summerscales2008maximum}, maximum
likelihood~\cite{gursel1989near} and Tikhonov regularization
scheme~\cite{rakhmanov2006rank, hayama2007coherent}.

%
% introduce machine learning
% 
In recent years, the field of machine learning and its sub-field, deep
learning, have been rapidly developing and have shown great potential in many
scientific fields~\cite{krizhevsky2012imagenet, NIPS2014_5423,
simonyan2014very, chen2014semantic, zeiler2014visualizing, szegedy2015going}.
For example, deep learning has been successfully applied to fields including
medical diagnosis~\cite{kononenko2001machine}, object
detection~\cite{redmon2016you}, image
recognition/processing/generation~\cite{he2016deep, krizhevsky2012imagenet,
zhang2016colorful, karpathy2015deep}, and  language
processing~\cite{lample2016neural}. In \ac{GW} astronomy, deep learning has
mostly been applied to the identifications of detector noise artefacts
(glitches)~\cite{mukund2017transient, zevin2017gravity, george2017deep} and the
detection of astrophysical signals~\cite{george2018deep, gabbard2018matching,
astone2018new} and their parameter estimation~\cm{add ref to Hunters latest
paper}. A \ac{CNN} is a deep learning algorithm that has the advantage of
capturing spatial and temporal features of the input data~\cm{this doesn't
really mean anything. Spatial refers to features in certain locations in images
and temporal refers to locations in a time series of in the time axis of a
spectrogram image}. Another advantage of the use of a \ac{CNN} in the detection
of a \ac{GW} signal is that a \ac{CNN} is relatively computationally cheap
compared to other more traditional methods. This is because the heavy
computational work is usually done during the training stage of a \ac{CNN}
prior to its actual application~\cite{goodfellow2016deep}.

%
% the main point of the paper
%
In this work, we demonstrate that a \ac{CNN} can be applied to the detection of
\acp{GW} from \acp{CCSN} and to the classification of their explosion
mechanisms.  To train our \ac{CNN}, we use simulated \ac{CCSN} waveforms from a
number of studies and simlate sources at a range of distances from $10$ to
$200$ kpc for two networks of four \ac{GW} detectors. The first network
consists of \ac{aLIGO}, \ac{AdVirgo} and KAGRA. For the second network, we
still include the detectors of \ac{AdVirgo} and KAGRA, but replace the two
detectors of \ac{aLIGO} with a modest set of planned upgrade version of them -
LIGO A+ in Hanford and Livingston~\cite{miller2015prospects, LIGOW}. We then
apply the \ac{CNN} to waveforms excluded from the training set to show that the
\ac{CNN} is not only able to identify the waveforms it familiarised itself with
during training, but also new waveforms~\cm{this is a strange statement. The
ability of the CNN to work on the training data is completely outweighed by its
perfromance on the independent testing set. The testing set perfroamnce is all
that you should mention here.}.

%
% paper structure
%
The remaining of this paper is structured as follows. In Section~\ref{sec:CNN},
we present a brief explanation of the concept of a \ac{CNN} as well as the
\ac{CNN} architecture  used for this study. In Section~\ref{sec:spwf}, we
introduce the waveforms we use for the training of the \ac{CNN} and  also
discuss the procedure with which we generated the input data for the \ac{CNN}.
The results are shown in Sections~\ref{sec:result} and~\ref{sec:unseen},
followed by our conclusions in Section~\ref{sec:conclusion}.
%\section{Supernova waveform}\label{sec:spwf}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Convolutional Neural network}\label{sec:CNN}

%
% describe the components of a CNN
%
A \ac{CNN} is a computational processing system that is composed of
interconnected layers of computational nodes~\cite{o2015introduction}. The
nodes are known as neurons and the outputs of which are processed with an
activation function. The activation function performs an elementwise nonlinear
operation to the output of the layer. In a standard \ac{CNN}, there are
commonly three types of layers: convolutional layers, max-pooling layers, and
fully connected layers~\cite{o2015introduction}. Convolutional layers perform
the mathematical operation of convolution between the weights of the layer's
neurons and the input to that layer. Max-pooling layers perform a down-sampling
process that reduces the size of the data by selecting the maximum of data
samples within fixed size bins. It can reduce the computational cost by
decreasing the number of trainable parameters (weights and biases) of the
\ac{CNN}.  Fully connected layers are layers that connect every neuron in its
layer to every neuron of its immediate previous and next layer. In the case of
detection and classification, the output layer is also a fully connected layer
with a softmax activation function with output used to represent the class
scores or probabilities.~\cm{even after editing, this isn't the clearest
description of a CNN. You are lacking an initial statement that says what a CNN
does - it takes in data and is able to classify it as one of the N types it has
learnt through training.}

%
% decribe the structure of the CNN
%
When multiple layers of these three types are stacked and connected one after
the other, a \ac{CNN} has been formed, where the output of each layer is the
input of the next layer. How these layers are connected in a \ac{CNN}, the
numbers of layers, their type, the number of neurons, the convolutional filter
size, max-pooling size, and the activation functions used, describes the
architecture of the \ac{CNN} (also known as hyper-parameters). The first layer
in a \ac{CNN}, also known as the input layer, is often a convolutional layer,
while the last layer or the output layer is often a fully connected layer with
an associated loss function~\cm{the loss isn't really associated with the
output layer. It is associated withe whole network.}. The architecture of the
remainder of the \ac{CNN} should depend on the specific task that it is being
trained to solve. An over-complicated model with too many trainable parameters
makes it easier to result in over-fitting (a procedure where the network
essentially memorises the training data but is unable to generalise to new
data). Alternatively, an overly simple architecture will struggle to capture
the features inherent to the input and not perform well in classification.
Automated schemes do exist for finding the optimal combination of
hyperparameters but in general (and in our case) the final architecture is
often obtained through rough trial and error, followed by fine tunning.

%
% describe training
%
During the training stage, the weights of the neurons in a \ac{CNN} are updated
using an algorithm called back-propagation~\cite{lecun1988theoretical}. The
output of the \ac{CNN} is used as an input to the loss-function~\cm{you need to
define what the loss function is supposed to do and how by minimising it you
make the network produce the right result. It isn't clear here.} and
back-propagation then computes the gradient of the loss-function with respect
to the traininable weights and biases within the network. A gradient descent
algorithm is then used to adjust the values of the weights and biases of the
neurons in each layer in order to iteratively minimise the loss-function.  When
the loss function is minimised and training is therefore complete, the \ac{CNN}
will be able to take in input in the form of new data and its output will best
represent the probability of that data belonging to each of the trained
classes. A well performing network will give high output probabilities to the
correct class in most cases. The process of achieving the minimisation of the
loss function during the training stage is the process whereby the machine is
"learning"..

%
% describe the specific architecture used here
% 
In this work, we employ a \ac{CNN} of $8$ convolutional layers, $3$ max-pooling
layers, and $3$ fully connected layers. The exact architecture of the \ac{CNN}
is shown in Table~\ref{table:architecture} and illustrated in
Fig.~\ref{fig:CNN}.
